Are populations like a circuit? The relationship between isolation by distance and isolation by resistance

### 2018-08-08 isolation_by_coalescence.tex
some figs
edited mostly until discussion

@@

\begin{figure}
\centering
    % \includegraphics[scale=.5]{conceptn}
    \caption{
    Illustration of the conceptual differences
    betweeen coalescence and commute time between states 1 and 3 
    for a continuous time Markov chain with four linearly connected states. 
    Coalescence time (left) is the time for two independently moving particles to meet and coalesce 
    (it is possible for them to be in the same location without coalescing).
    Commute time is time for a single particle starting in one state to travel to another state 
    and then back to the original state.
    Note how the time axis is scaled to be faster in the figure showing commute time
    than the one for coalescence time in order for them to fit in the same space.
    } \label{fig:concept_coalcom}
\end{figure}

@@

\begin{figure}
\centering
% \includegraphics[scale=.7]{valid_range}
\caption{The green shaded region shows 
	the valid range of the parameter values
	for the two state Markov chain given that
	$C_{11}=1$, $C_{12}=2$, and $C_{22}=1.5$}
\label{fig:valid_range}
\end{figure}

@@

\paragraph{MCMC methods}
To sample from the posterior distribution of equation \eqref{eq:post}, 
we used R \citep{R} 
to implement a standard Metropolis-Hastings algorithm with Gaussian proposals \citep{brooks2011handbook}.
Starting locations were chosen by sampling from the prior distribution.
Before the usual ``burn-in'' phase before sampling for inference begins,
we included a period of ``pre-burn-in'' which used the same MCMC procedure
with a larger $\sigma_\epsilon$, to allow the chain to more quickly converge on the high-posterior
portion of parameter space.
Typically, we ran MCMC for $10^6$ iterations of pre-burn-in 
and $3 \time 10^6$ iterations of burn-in,
followed by $4 \times 10^6$ iterations which were used to estimate posterior distributions.
We used standard methods to assess convergence and mixing.

@@

\plr{transition paragraph}
% biased migration is bad for commute times: 
%              figure 19 showing the two against each other, and the simple example
% continuous space: can find big barriers but lots of between-replicate variability (life is still hard)
% tortoises: results on small graph, showing strong asymmetry (?)

\plr{Summarizing observations about inference being harder with more noise,
more parameters,
and lower coalescence rate.}

@@

\plr{TODO: replace bottom two panels in figure 18 with same thing for commute time}

\plr{pasted in}
We now test the method's ability to locate barriers to migration
on a $5 \times 3$ grid with asymmetric gene flow. 
There are impassable boundaries on the bottom left and top right of the landscape,
but the overall graph remains fairly well connected (see Figure \ref{fig:5x3b_grid}).
Each value of $g$ in the true model is determined randomly as before.
The coalescence rates are set to 1 everywhere, 
and for the first case, we assume that we know it is the same everywhere, 
although the specific value is not known.
Gaussian noise with standard deviation equal to 1/1000 of the mean of $H$ is added to form $\hat{H}$, 
and inference is performed using all values of $\hat{H}$.
All inference in this section is done using
$5 \times 10^6$ iterations of MCMC 
after $4 \times 10^6$ iterations for burn-in and $1 \times 10^6$ iterations for pre-burn-in.
In this case, the mean absolute error of $\hat{g}$ was 0.019, 
while the error of $\hat{\gamma}$ was 0.0028. %\emph{talk more about results here?}
This analysis is repeated in three more situations, 
all with noise set to have standard deviation 1/100 of the mean of $H$.
In the first case, there are no missing values of $\hat{H}$, 
in the second case, values from rows and columns
corresponding to populations 2, 5, 7, and 15 are missing from $\hat{H}$. 
The third case is the same as the second except that
there is a coalescence parameter for each location on the grid 
instead of one parameter for all locations.
With no missing values and noise of 1/100, 
the mean absolute error of $\hat{g}$ was 0.17 and the error of $\hat{\gamma}$ was 0.0066.
With the four missing values,
the mean absolute error of $\hat{g}$ was 0.37 and the error of $\hat{\gamma}$ was 0.033.
With missing values and multiple coalescence parameters, 
the mean absolute error of $\hat{g}$ was 0.52 
and the mean absolute error of $\hat{\gamma}$ was 0.42.
Boxplots of the posterior distributions of $g$ can be seen in Figure \ref{fig:5x3boxplots}.
We can see that the inference method does very well 
in the case with noise of 1/1000.
With each modification, the problem becomes less well defined, 
so it is not surprising that the error increases as well.
Notably, we are still able to identify the barrier 
except when there is data missing from both locations on either side of the barrier, 
as is the case for movement parameters $g_{5}$ and $g_{17}$ with locations 2 and 7.
Additionally, inference of the coalescence rate is very accurate 
when there is only one coalescence rate parameter,
but is much harder with multiple parameters, 
which in turn makes the movement parameters harder to estimate.

%This is repeated with noise set to have standard deviation 1/100 of the mean of $H$ 
%and with data missing from 4 of the 15 locations (\emph{also non-constant coalescence?}).
%In this case, the mean absolute error of $\hat{g}$ was (insert) and (appropriate statement for $\gamma$)
%after (insert) iterations of MCMC.
%\emph{talk more about results here}

\begin{figure}
\centering
% \includegraphics[scale=.6]{5x3b_grid}
\caption{Plot of the grid structure 
for the 5 $\times$ 3 grid with barriers.
Values for the non-zero movement rates 
were chosen by rounding independent draws of an exponential random variable with mean 1 
up to the nearest tenth.}
\label{fig:5x3b_grid}
\end{figure}

\begin{figure}
\centering
% \includegraphics[scale=0.8]{5x3boxplots}
\caption{Posterior distributions for values of $g$ 
for the 5 $\times$ 3 graph with barriers 
for each of the analysis cases.
See Figure \ref{fig:5x3b_grid} to see movement parameter locations.}
\label{fig:5x3boxplots}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Biased migration}
\label{sec:comvcoal}

\plr{transition about how we expect biased migration to produce large differences}

In order to further illustrate the differences between commute time and coalescence time inference,
we test both methods on four different $4 \times 4$ graphs (shown in Figure \ref{fig:4x4_grids}:
one where movement parameters are all equal to 1 (which we will call the uniform case),
one where movement parameters are symmetric but randomly generated, 
one asymmetric graph where all movement parameters are randomly generated,
and one asymmetric graph where there is a bias in moving towards one direction in the graph.
For the symmetric and unbiased asymmetric graphs, movement parameters are chosen randomly as before.
In the biased asymmetric graph, all parameters for moving left or down are 2 while 
all parameters moving right or up are 0.5.
Differences between true mean coalescence times ($H$) and mean commute times ($\comdist$) 
are shown for each graph in Figure \ref{fig:4x4coalvcomH}. 
The differences are fairly small in the uniform and symmetric cases, moderate in the asymmetric case, 
and extreme in the biased asymmetric case.
%The structure of the graphs themselves can be seen in Figure \ref{fig:4x4_grids}.
Inference is done using $\hat{H}$ with a noise level of 1/200 the mean value of $H$ for each grid.
All trials with each inference method were done using $4 \times 10^6$ iterations of MCMC 
after $1 \times 10^5$  iterations of pre-burn-in and $1 \times 10^6$ iterations of burn-in.
Posterior distributions for each graph and and inference method are shown in Figure \ref{fig:4x4box}.
The coalescence time inference was substantially more accurate in all cases.
The commute time based inference and coalescence time based inference methods had 
mean absolute errors in $\hat{g}$ of 0.30 and 0.12 respectively for the uniform graph,
0.49 and 0.053 for the symmetric graph, 
0.87 and 0.035 for the asymmetric graph,
and 3.27 and 0.053 for the biased asymmetric graph, respectively.
We can see that, as expected based on the differences between $H$ and $\comdist$,
the coalescence time inference performs very well in all four test cases
while the commute time inference performs fairly well in uniform and symmetric cases,
poorly in the asymmetric case, and very poorly in the biased asymmetric case.
The behavior of the biased asymmetric graph supports our previous thoughts about
situations similar to the three state system seen in Figure \ref{fig:3_state},
where commute times are much longer than coalescence times.
In this case, the commute time inference predicts that many movement parameters
are an order of magnitude larger than they are in truth. 
%and there does not appear to be a correlation between 
%the magnitude of the true and inferred values, 
%indicating that we cannot infer much beyond that the population mixes quickly,
%which was already clear from the data itself.

%\emph{talk more about results here}

\begin{figure}
\centering
% \includegraphics[scale=.8]{4x4_grids}
\caption{Plots of the grid structure 
for the uniform graph, symmetric graph, the asymmetric graph,
and the biased asymmetric graph 
using the igraph R package.}
\label{fig:4x4_grids}
\end{figure}

\begin{figure}
\centering
% \includegraphics[scale=.8]{4x4coalvcomH}
\caption{Plots of the true values of $\comdist$ vs $H$ 
for the uniform graph, symmetric graph, the asymmetric graph,
and the biased asymmetric graph, 
shown with the $y=x$ line.
See Figure \ref{fig:4x4_grids} to see overall graph structure.}
\label{fig:4x4coalvcomH}
\end{figure}

\begin{figure}
\centering
% \includegraphics[scale=.8]{4x4box}
\caption{Boxplots of the posterior distributions for $g$ for commute time and coalescence time
based inference for the uniform graph, symmetric graph, 
the asymmetric graph, and the biased symmetric graph.
See Figure \ref{fig:4x4_grids} to see movement parameter locations.}
\label{fig:4x4box}
\end{figure}

@@

\plr{transition about discretization and continuous space}

In order to validate the methods in situations 
that are more similar to those that we would see with actual populations,
we performed simulations using SLiM (Selection on Linked Mutations) \citep{haller2016slim}.

\plr{move to methods}
We used SLiM to simulate discrete generations based on the Wright-Fisher model.
Simulations were performed in three different situations:
uniform migration on a continuous square landscape, 
uniform migration on a rectangular landscape with barriers to migration, 
and biased migration on a square landscape.
The simulations performed in this section 
were of populations of 10000 diploid hermaphroditic individuals.
Mates were selected from nearby individuals using a Gaussian kernel 
with standard deviation of $\sigma_{d}$,
truncated at $3 \sigma_{d}$.
In each simulation,
$\sigma_{d}$ was chosen such that there would be an average of 
0.25 individuals per $\sigma_{d}^2$ (2.25 per $(3 \sigma_{d})^2$),
which is 0.005 for a square landscape with one unit of area.
Offspring are dispersed according to a Gaussian distribution away from the mother
using the same truncated Gaussian kernel as before.
If the location of the offspring falls outside of the habitable area, 
it is redrawn until it is in a valid location.
In order to prevent excessive clumpiness of the population density distribution \citep{felsensten1975pain}, 
we also model local competition. 
To do this, we use the same truncated Gaussian kernel,
with the strength of the interaction chosen 
by visual inspection of the density distribution.
Specifically, the new fitness of each individual after accounting for local competition
is equal to $\exp{(-\text{totalStrength}/40)}$,
where totalStrength is the sum of the strength of interations with all individuals
within the cutoff distance according to the Gaussian kernel,
resulting in reduced fitness for individuals in dense regions.
Individuals are chosen to reproduce proportional to their fitness until the next generation.
Clumpiness is similar to what is seen in 
Figure \ref{fig:ind_locs_5x3b_1}

In our simulations, individuals have 10 pairs of chromosomes with 200,000 base pairs each 
with a recombination rate of $10^{-8}$ per generation per base pair.
In the landscapes with barriers and biased migration,
mutation rate was set to $10^{-7}$.
In order to speed up simulation in the case of the landscape with uniform migration,
the mutation rate was lowered to $10^{-8}$.
This should not have a dramatic effect on the pattern of genetic relatedness between individuals
aside from reducing the genetic distance for those simulations by a factor of ten.

In each location, individuals are sampled uniformly from those 
within the middle three quarters in each dimension of the grid space,
as shown for one situation iin Figure \ref{fig:ind_locs_5x3b_1}.
This protocol was chosen as a compromise between sampling all individuals 
from near the center of each location,
which could result in a large number of close relatives being sampled, 
leading to an underestimate of the mean genetic distance for individuals in that location
and uniform sampling from the whole area of each grid, 
which may result in many sampled individuals being very close to other grid squares.
An equal number of individuals are drawn from each location,
but this need not be the case.
We could even have some locations from which no individuals were drawn,
but this lowers performance as discussed in Section \ref{sec:5x3b}.
Mean genetic distances for each pair of locations 
are calculated from the means of all of the pairwise genetic divergences 
between genomes in the two locations, excluding self comparisons.
Standard error of the mean is calculated conservatively using 
the minumum number of independent samples, 
in this case the smallest number of individuals from each location.

Before being passed to the MCMC inference method,
genetic distance values are scaled so that the resulting parameters 
are approximately of order one.
This is accomplished by multiplying the mean genetic distances by a constant
such that the new mean of the mean genetic distances 
is equal to the number of states in the discretization. 
For example, if we discretize a landscape into 16 squares, 
the mean of the mean genetic distances after scaling will be 16.

For the $4 \times 4$ discretizations of the uniform and biased migration cases, 
as well the barrier case, 
inference was performed using $4 \times 10^6$ iterations of MCMC
after $1 \times 10^6$ iterations of pre-burn-in and $3 \times 10^6$ iterations of burn-in. 
For the $2 \times 2$ discretizations, $1 \times 10^6$ iterations were used 
with $1 \times 10^5$ for pre-burn-in and $1 \times 10^6$ for burn-in.

% {Continuous Landscape with a Barrier}
We now revisit the earlier situation where a landscape has locations that are barriers to gene flow,
as in Section \ref{sec:5x3b}.
The landscape is designed to be a continuous version of this situation.
Since the distribution of offspring locations is truncated at three standard deviations
away from the mother, a sufficiently thick uninhabitable area can completely stop migration
directly across it.
As in \ref{sec:5x3b}, though, the landscape is still relatively well connected.
The layout of the landscape can be seen in Figure \ref{fig:ind_locs_5x3b_1},
with the red bars being the uninhabitable regions that serve as barriers to migration.
Unlike the case of a square landscape with no barriers, 
there is not a natual way to discretize it more coarsely 
with grid squares of equal size and shape,
so a $5 \times 3$ grid will be the only discretization used in this section.
We calculate mean genetic distances using 50 individuals 
from each grid location,
here giving us a total of 750 individuals.

\begin{figure}
\centering
% \includegraphics[scale=.8]{ind_locs_5x3b_1}
\caption{Locations of all and selected individuals on the $5 \times 3$ landscape with barriers
in the first replicate.
Locations of individuals used to compute the mean genetic distance matrix are shown in bold points
whereas the locations of other individuals are shown in smaller points.
The red bars show the locations of uninhabitable regions of the landscape.
They are sufficiently thick so that no migration can occur directly across them.}
\label{fig:ind_locs_5x3b_1}
\end{figure}

\begin{figure}
\centering
% \includegraphics[scale=.8]{grid_5x3b_1}
\caption{Medians of the posterior distributions of the inferred values of $g$ 
for the first replicate of the $5 \times 3$ landscape with barriers.
The gene flow rate across the barriers are small, 
but there is no clear pattern amoung the rest of the values.}
\label{fig:grid_5x3b_1}
\end{figure}

\begin{figure}
\centering
% \includegraphics[scale=.8]{post_dists_5x3b_1_new}
\caption{Boxplots of the posterior distributions for $g$ 
for the first replicate of the $5 \times 3$ landscape with barriers.
Boxes for gene flow rates for edges that cross the barrier are colored in red.}
\label{fig:post_dists_5x3b_1}
\end{figure}

@@

The third situation that we will discuss is where migration is biased.
We will be using a square landscape as in the first case, 
but the offspring locations of offspring are not centered around the parent.
Instead, the mean of the distribution of the locations
is one tenth of a standand deviation (or 0.05\% of the edge length of the landscape) 
up and to the right of their parent's location.
This results in reverse time gene flow down and to the left 
as in the biased asymmetric graph in Section \ref{sec:comvcoal}.
In terms of population genetics, 
this is very similar to the case where a population is expanding into new territory 
since in both cases, 
offspring tend to be further from the edge of the expansion than their parents.
The difference is that, in our case, 
the population is constrained to be a bounded distance 
away from the edge of the ``expansion''.
In both situations, individuals at the edge may have disproportionately more offspring,
leading to low genetic diversity and strong drift.
This results in substantially more noise in the isolation by distance plots 
as seen in Figure \ref{fig:ibd_comp}. 
The horizontal lines in that figure are likely from comparing genomes of individuals from
two families that have recently increased in size and geographical spread, 
resulting in many pairs of genomes with similar genetic distance 
but varying geographic distance.

@@

\begin{figure}
\centering
% \includegraphics[scale=.35]{ibd_comp}
\caption{Comparison of the isolation by distance plots for the square landscape with uniform migration
and the square landscape with biased migration.
Faster coalescence in the biased migration case results in more noticeable horizontal lines,
indicative of families that recently increased in size and geographical spread.}
\label{fig:ibd_comp}
\end{figure}

\begin{figure}
\centering
% \includegraphics[scale=.5]{2x2grid_asb}
\caption{Medians of the posterior distributions of the inferred values of $g$ 
for the first replicate of the $2 \times 2$ discretization of the square landscape with biased migration.
The reverse time gene flow rates out of the bottom left section are very small,
indicating that very few individuals living in the bottom left of the landscape
have ancestors from other regions of the habitat,
consistent with migration biased up and to the right.}
\label{fig:2x2grid_asb}
\end{figure}

\begin{figure}
\centering
% \includegraphics[scale=.8]{grid_bias_4x4_1}
\caption{Medians of the posterior distributions of the inferred values of $g$ 
for the first replicate of the $4 \times 4$ discretization of the square landscape with biased migration.
As in the $2 \times 2$ case, 
the reverse time gene flow rates out of the bottom left section are very small,
indicating that very few individuals living in that area of the landscape
have ancestors from other regions of the graph.
Most, but not all of the other gene flow rates are biased 
in the direction we would expect.}
\label{fig:grid_bias_4x4_1}
\end{figure}

\begin{figure}
\centering
% \includegraphics[scale=.6]{post_dists_bias_4x4_1}
\caption{Boxplots of the posterior distributions for $g$ 
for the first replicate of the $4 \times 4$ discretization of the square landscape with biased migration.
Black boxes are gene flow rates down and to the left 
with each green box being the corresponding gene flow rate up and to the right.}
\label{fig:post_dists_bias_4x4_1}
\end{figure}

\section*{Application of Inference to Mojave Desert Tortoises}

We also apply the method to a population of Mojave desert tortoises (\emph{Gopherus agassizii})
living in California. 
Location and low coverage whole genome sequencing data was collected from 271 individuals 
\citep{shaffer2017desert}.
Pairwise genetic distances between all individuals was obtained from the same study.
The landscape was discretized into 13 regions based on watersheds and the locations of individuals.
The mean coalescence time matrix and standard deviations were calculated as before. 
Since the regions vary substantially in size, 
and may vary substantially in population density, 
we do not assume that coalescence rates are the same everywhere, 
which gives 13 values of $\gamma$. 
Since allowing only nearest neighbor migration gives 50 values of $g$, 
we have a total of 63 unknown parameters.
Because there are $13(14)/2 = 91$ unique values of the mean coalescence time matrix, 
we have more equations than unknowns.

>>>>

We apply the method to the desert tortoise \plr{say more here}.

@@

\begin{figure}
\centering
     \includegraphics[scale=.5]{figs/conceptn}
    \caption{
    Illustration of the conceptual differences
    betweeen coalescence and commute time between states 1 and 3 
    for a continuous time Markov chain with four linearly connected states. 
    Coalescence time (left) is the time for two independently moving particles to meet and coalesce 
    (it is possible for them to be in the same location without coalescing).
    Commute time is time for a single particle starting in one state to travel to another state 
    and then back to the original state.
    Note how the time axis is scaled to be faster in the figure showing commute time
    than the one for coalescence time in order for them to fit in the same space.
    } \label{fig:concept_coalcom}
\end{figure}

@@

\begin{figure}
\centering
         \includegraphics[scale=.7]{figs/valid_range}
    \caption{The green shaded region shows 
        the valid range of the parameter values
        for the two state Markov chain given that
        $C_{11}=1$, $C_{12}=2$, and $C_{22}=1.5$}
    \label{fig:valid_range}
\end{figure}

@@

\paragraph{MCMC methods}
To sample from the posterior distribution of equation \eqref{eq:post}, 
we used R \citep{R} 
to implement a standard Metropolis-Hastings algorithm with Gaussian proposals \citep{brooks2011handbook}.
Starting locations were chosen by sampling from the prior distribution.
Before the usual ``burn-in'' phase before sampling for inference begins,
we included a period of ``pre-burn-in'' which used the same MCMC procedure
with a larger $\sigma_\epsilon$, to allow the chain to more quickly converge on the high-posterior
portion of parameter space.
Typically, we ran MCMC for $10^6$ iterations of pre-burn-in 
and $3 \time 10^6$ iterations of burn-in,
followed by $4 \times 10^6$ iterations which were used to estimate posterior distributions,
although for small graphs fewer iterations were necessary.
We used standard methods to assess convergence and mixing.

@@

\plr{transition paragraph}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Model validation}

We first tested the method using data generated under the model.
To test the methods across a wide range of situations,
we generated random movement rates for 
a large number of $3 \times 2$ rectangular grids:
twenty-five symmetric (i.e., with $G_{ij} = G_{ji}$) and twenty-five asymmetric.
We generated movement parameters
by rounding independent exponential random variables up to the nearest one tenth,
and fixed coalescence rate at 1 for all locations.
We then added independent noise to each matrix entry (preserving symmetry),
and assume that we can estimate the amount of noise reasonably well, 
so used the standard deviation of the noise for $\sigma_\epsilon$ in the likelihood function.
To assess the impact on inference of the number of parameters,
we performed inference both with and without the assumption 
that coalescence rate is the same everywhere.
For $3 \times 2$ graphs, there are 21 equations
and 14 movement parameters,
so that with a global coalescence rate there are 15 unknowns
but with individual coalescence rates there are 20 unknowns,
which is close to nonidentifiable.
(For commute time inference, we analogously constrained all values of $q$ to be identical or not.)
To quantify accuracy across these random graphs in different situations,
we computed, for each model fit on a particular random graph,
the absolute difference between the posterior median of each movement parameter
and the true value.
We then averaged these across movement parameters to produce a measure of accuracy for that model fit,
reported below as ``mean absolute error''.

\paragraph{Varying noise}
Unsurprisingly, coalescence time inference became less accurate as the amount of noise increased.
As we varied the standard deviation of the noise
from 1/1000 to 1/50 of the mean value of $C$ for that grid,
the mean absolute error 
for coalescence time inference with a single coalescence rate
increased from around 0.05 to 0.4.
Posterior interquartile ranges increased proportionally.
(although there was substantial variation in accuracy between replicates;
see Figures \ref{fig:mult_noise} and \ref{fig:mult_noise_iqr}).
Increasing the number of parameters by allowing multiple coalescence rates
made the inference problem even harder --
mean absolute errors no longer depended strongly on the amount of noise added,
and hovered around 0.4.

More surprisingly, the performance of commute time inference did not depend on the amount of noise --
mean absolute error was around 0.6 across all levels of noise,
both with and without more than one parameter for local diversity
(see Figure \ref{fig:mult_noise}).
Although symmetry did not affect coalescence time inference,
commute time inference did substantially worse on asymmetric graphs,
showing median absolute errors of around 0.8.

\paragraph{Varying coalescence rates}
Lower coalescence rates mean that lineages wander about the landscape for longer
before they coalesce, 
thus greatly reducing the geographic information we get from coalescence times.
Indeed, coalescence rates had a strong effect on feasability of the inference problem
(shown in Figures \ref{fig:mult_gam} and \ref{fig:mult_gam_iqr}):
lowering the coalescence rate from 10 to 0.1
(always adding noise with standard deviation 1/200th of the mean value)
increased mean absolute error of coalescence time inference from around 0.05 to 0.5.
Again, allowing multiple coalescence rates
or using commute time inference were highly inaccurate in all cases,
with mean absolute errors of above 0.5.

@@

We then test the method's ability to locate barriers to migration
on the $5 \times 3$ grid with asymmetric gene flow shown in Figure \ref{fig:5x3b_grid}. 
Each nonzero movement rate is determined randomly as before.
Coalescence rates are set to 1 everywhere.
Figure \ref{fig:5x3boxplots} shows posterior distributions of the movement rates
for both coalescence and commute time inference
on this graph, with noise standard deviation equal to
1/1000 of the mean coalescence time (left)
and 1/100 of the mean coalescence time (right).
Although increased noise in the data increased uncertainty in the estimated values,
coalescence time inference correctly inferred the location of the barrier
(as well as high-dispersal connections) in both cases.
Commute time inference \plr{did what?}
We also explored making the problem harder,
by first (a) dropping rows and columns from the data corresponding to populations 2, 5, 7, and 15;
and then (b) allowing a separate coalescence rate for each location,
in both cases using a noise level of 1/100.
The resulting posterior distributions are shown in Figure \ref{fig:5x3boxplots_mult_gam},
and show substantially increased error,
but only in movement rates to the removed locations.
As before, uncertainty was increased when multiple coalescence rates were allowed,
but not as strongly as before, likely because the problem is less close to being underconstrained,
having 105 equations and 59 unknowns.

\begin{figure}
\centering
     \includegraphics[scale=.6]{figs/5x3b_grid}
    \caption{Plot of the grid structure 
    for the 5 $\times$ 3 grid with two barriers:
    movement rates between $\{2,3\}$ and $\{7,8\}$ are zero,
    as are those between $\{8,9\}$ and $\{13,14\}$.
    Values for the non-zero movement rates 
    were chosen by rounding independent draws of an exponential random variable with mean 1 
    up to the nearest tenth.}
    \label{fig:5x3b_grid}
\end{figure}

\begin{figure}
\centering
     \includegraphics[scale=0.8]{figs/5x3boxplots}
    \caption{Posterior distributions for values of $g$ 
    for the 5 $\times$ 3 graph with barriers 
    for each of the analysis cases.
    See Figure \ref{fig:5x3b_grid} to see movement parameter locations.
    \plr{TODO: replace bottom two panels with same thing for commute time}
}
    \label{fig:5x3boxplots}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Biased migration}
\label{sec:biased_migration}

Both matrices of coalescence and commute times, $C$ and $R$, are symmetric,
but they do not deal with asymmetric (biased) migration in the same way.
Since commute time to $i$ requires paths both to and from $i$,
but coalescence time to $i$ only reaquires lineages to leave $i$,
if, for instance, there is a low rate of migration back into $i$,
then commute times to $i$ could be quite long even while coalescence times are not.
For instance, suppose there are three populations arranged in a line,
and lineages currently in the outer populations are much more likely to come from the central island
than the other way around.
Since lineages quickly move to the center and coalesce regardless of starting position,
coalescence time (and genetic diversity) will be relatively low between all individuals.
Commute time between the two outer locations, on the other hand,
will be much longer than other comparisons, since it requires a lineage to leave the center.

In order to investigate this general situation,
we tested both methods on four different $4 \times 4$ graphs (shown in Figure \ref{fig:4x4_grids}):
\emph{uniform}, where all movement rates are 1.0;
\emph{symmetric}, where movement rates are symmetric and randomly generated as before;
\emph{asymmetric}, where all movement rates are random;
and \emph{biased}, where movement rates down or to the left are equal to 2.0,
and movement rates up or to the right are 0.5.

Differences between true mean coalescence times ($C$) and mean commute times ($\comdist$) 
are shown in Figure \ref{fig:4x4coalvcomH}:
they are fairly small in the uniform and symmetric cases, moderate in the asymmetric case, 
and extreme in the biased asymmetric case.
This difference suggests that commute time inference may be strongly misled 
in the asymmetric and biased situations,
but it still could be the case that the movement rates that give the best fit of commute time
to coalescence time data could be close to the actual rates use to generate the data.

Nonetheless, 
coalescence time inference was substantially more accurate in all cases.
Posterior distributions of movement parameters are shown in Figure \ref{fig:4x4box},
produced by applying both inference methods
after adding noise with standard deviation of 1/200 the mean value of $C$.
The mean absolute errors for coalescence time inference are 0.12 or less in all cases.
Commute time inference obtains roughly uniform migration rates
(although varying by a factor of 2) in the uniform case;
and migration rates noisily correlated with the truth in the symmetric graph.
However, commute time inferences for the asymmetric and biased graphs are only weakly correlated, 
if at all, to the truth.
Commute times also drastically overestimate movement rates for the biased case,
as we would expect based on the differences discussed above.

\begin{figure}
\centering
     \includegraphics[scale=.8]{figs/4x4coalvcomH}
    \caption{Plots of the true values of $\comdist$ vs $H$ 
    for the uniform graph, symmetric graph, the asymmetric graph,
    and the biased asymmetric graph, 
    shown with the $y=x$ line.
    See Figure \ref{fig:4x4_grids} to see overall graph structure.}
    \label{fig:4x4coalvcomH}
\end{figure}

\begin{figure}
\centering
     \includegraphics[scale=.8]{figs/4x4box}
    \caption{Boxplots of the posterior distributions for $g$ for commute time and coalescence time
    based inference for the uniform graph, symmetric graph, 
    the asymmetric graph, and the biased symmetric graph.
    See Figure \ref{fig:4x4_grids} to see movement parameter locations.}
    \label{fig:4x4box}
\end{figure}

@@

The data we have used thus far are produced ``under the model'',
and should provide an accurate depiction of our method applied to discrete, randomly mating populations
whose connections by migration are known.
However, this is nearly always a rough approximation to reality,
in which organisms are distributed across continuous geography,
and ``populations'' are constructed by necessity, often driven by sampling locations.

\paragraph{Process noise}
Talk about Figure \ref{fig:h_val_comp}.

\begin{figure}
\centering
\includegraphics[scale=.8]{figs/h_val_comp}
\caption{
    Comparison of the values of the genetic distance matrix $C$
    for a $4 \times 4$ discretization of a square landscape with uniform migration
    calculated from various situations.
    All values are plotted against the corresponding values from the $C$ matrix
    calculated using the first set of individuals from the first simulation,
    along with the $y=x$ line.
    Black circles are from a second (non-overlapping) set of individuals from the first simulation,
    red triangles are from the second simulation, 
    and blue plusses are for the third simulation.
    Mean absolute value relative differences between the new values of $C$ for each situation
    and the original values from the first set of individuals from the first simulation are
    1.8\% for the second set of indivuals from the first simulation,
    13.2\% for the second simulation, 
    and 7.3\% for the third simulation.
    } \label{fig:h_val_comp}
\end{figure}

\paragraph{Identifying a barrier}
We now revisit the earlier situation where a landscape has locations that are barriers to gene flow,
as in Section \ref{sec:5x3b}.
The landscape is designed to be a continuous version of this situation.
Since the distribution of offspring locations is truncated at three standard deviations
away from the mother, a sufficiently thick uninhabitable area can completely stop migration
directly across it.
As in \ref{sec:5x3b}, though, the landscape is still relatively well connected.
The layout of the landscape can be seen in Figure \ref{fig:ind_locs_5x3b_1},
with the red bars being the uninhabitable regions that serve as barriers to migration.
Unlike the case of a square landscape with no barriers, 
there is not a natual way to discretize it more coarsely 
with grid squares of equal size and shape,
so a $5 \times 3$ grid will be the only discretization used in this section.
We calculate mean genetic distances using 50 individuals 
from each grid location,
here giving us a total of 750 individuals.

\begin{figure}
\centering
     \includegraphics[scale=.8]{figs/ind_locs_5x3b_1}
    \caption{Locations of all and selected individuals on the $5 \times 3$ landscape with barriers
    in the first replicate.
    Locations of individuals used to compute the mean genetic distance matrix are shown in bold points
    whereas the locations of other individuals are shown in smaller points.
    The red bars show the locations of uninhabitable regions of the landscape.
    They are sufficiently thick so that no migration can occur directly across them.}
    \label{fig:ind_locs_5x3b_1}
\end{figure}

\begin{figure}
\centering
     \includegraphics[scale=.8]{figs/post_dists_5x3b_1_new}
    \caption{Boxplots of the posterior distributions for $g$ 
    for the first replicate of the $5 \times 3$ landscape with barriers.
    Boxes for gene flow rates for edges that cross the barrier are colored in red.}
    \label{fig:post_dists_5x3b_1}
\end{figure}

@@

The third situation that we will discuss is where migration is biased.
We will be using a square landscape as in the first case, 
but the offspring locations of offspring are not centered around the parent.
Instead, the mean of the distribution of the locations
is one tenth of a standand deviation (or 0.05\% of the edge length of the landscape) 
up and to the right of their parent's location.
This results in reverse time gene flow down and to the left 
as in the biased asymmetric graph in Section \ref{sec:biased_migration}.
In terms of population genetics, 
this is very similar to the case where a population is expanding into new territory 
since in both cases, 
offspring tend to be further from the edge of the expansion than their parents.
The difference is that, in our case, 
the population is constrained to be a bounded distance 
away from the edge of the ``expansion''.
In both situations, individuals at the edge may have disproportionately more offspring,
leading to low genetic diversity and strong drift.
This results in substantially more noise in the isolation by distance plots 
as seen in Figure \ref{fig:ibd_comp}. 
The horizontal lines in that figure are likely from comparing genomes of individuals from
two families that have recently increased in size and geographical spread, 
resulting in many pairs of genomes with similar genetic distance 
but varying geographic distance.

@@

\begin{figure}
\centering
 \includegraphics[scale=.35]{figs/ibd_comp}
\caption{Comparison of the isolation by distance plots for the square landscape with uniform migration
and the square landscape with biased migration.
Faster coalescence in the biased migration case results in more noticeable horizontal lines,
indicative of families that recently increased in size and geographical spread.}
\label{fig:ibd_comp}
\end{figure}

\begin{figure}
\centering
 \includegraphics[scale=.5]{figs/2x2grid_asb}
\caption{Medians of the posterior distributions of the inferred values of $g$ 
for the first replicate of the $2 \times 2$ discretization of the square landscape with biased migration.
The reverse time gene flow rates out of the bottom left section are very small,
indicating that very few individuals living in the bottom left of the landscape
have ancestors from other regions of the habitat,
consistent with migration biased up and to the right.}
\label{fig:2x2grid_asb}
\end{figure}

\begin{figure}
\centering
 \includegraphics[scale=.8]{figs/grid_bias_4x4_1}
\caption{Medians of the posterior distributions of the inferred values of $g$ 
for the first replicate of the $4 \times 4$ discretization of the square landscape with biased migration.
As in the $2 \times 2$ case, 
the reverse time gene flow rates out of the bottom left section are very small,
indicating that very few individuals living in that area of the landscape
have ancestors from other regions of the graph.
Most, but not all of the other gene flow rates are biased 
in the direction we would expect.}
\label{fig:grid_bias_4x4_1}
\end{figure}

\begin{figure}
\centering
 \includegraphics[scale=.6]{figs/post_dists_bias_4x4_1}
\caption{Boxplots of the posterior distributions for $g$ 
for the first replicate of the $4 \times 4$ discretization of the square landscape with biased migration.
Black boxes are gene flow rates down and to the left 
with each green box being the corresponding gene flow rate up and to the right.}
\label{fig:post_dists_bias_4x4_1}
\end{figure}

\section*{Application to the Mojave Desert Tortoise}

The Mojave desert tortoise (\textit{Gopherus agassizii})
lives across much of the Mojave desert of California and Nevada,
separated from its sister taxa, \textit{Gopherus morofkaii} by the Colorado river.
We applied our method to genomic distances calculated from
whole genome sequencing data of 271 individuals sampled from across much of the range,
reported in \citet{shaffer2017desert}.
We discretized the landscape into 13 regions based on watershed polygons
from the Watershed Boundary Dataset \citep{WBD}
(using WBD8 slightly modified by portions of WBD10 to improve contiguity of regions).
The resulting discretization,
overlaid on a map including tortoise sampling locations,
is shown in Figure \ref{fig:tort_land}.
Since the regions vary substantially in size, 
and may vary substantially in population density, 
we do not assume that coalescence rates are the same everywhere, 
which gives 13 coalescence rates;
combined with the 50 migration parameters between each pair of adjacent regions
we have a total of 63 unknown parameters
(and 91 equations).
